#################################
##
## GROUP 8, ASSIGNMENT #2, York CSML1000 Fall 2019 .Rmd Starter File with CRISP-DM.
##
## In this skeleton of a file, we outline the major steps of CRISP-DM, in the hopes of touching on as many
## of the points as possible. (Some may be changed or combined along the way, but this is the roadmap.)
##
## Quick link to Assignment Specification: https://github.com/3dspace/CSML1000-Assignment-2/blob/master/specification-assignment-2.pdf
##
## [ Delete this chunk when we're under way ]
##
#################################

---
title: "Clustering Credit Card Customers for Targeted Marketing using Unsupervised Learning"
output: html_document
---

# CSML1000 Project #2, by Group 8
Tamer Hanna <tamerh@my.yorku.ca>
Pete Gray <ptgray@my.yorku.ca>
Xiaohai Lu <yu271637@my.yorku.ca>
Haofeng Zhou <zhf85@my.yorku.ca>

```{r message=FALSE, warning=FALSE}
# Use R chunks like these, unless we determine otherwise
```

#OVERVIEW

Using data on the behaviour of credit card customers, we can use unsupervised learning to discover market segments that would be useful for targeting marketing strategies.

Our dataset has 19 columns of data on 9000 customers. Using K-means and PCA, we can determine an optimal number of market segments, discover the identifying properties of those segments, and be able to describe to the marketing department what the most significant behaviours of the people in those segments are. Marketing campaigns, then, can be targeted at, for example, impulse shoppers, big spenders, or people who have a hard time paying off their debts.

# BUSINESS UNDERSTANDING

Applying specific marketing strategies to different types of customers can improve results and reduce costs. Credit card customers can be profiled by the way they use, and pay off, their card.

## Business Objectives

Behavioural data can be mined to discover identifying features of clusters of customers who use their card in similar ways. The marketing department can use these insights to select target groups and optimize the marketing used to target them.
 
## Data Mining Goals

Using shallow algorithm unsupervised learning, we can discover clusters containing customers who exhibit similar behaviours. We can examine those 


# DATA UNDERSTANDING

## Collect Initial Data

```{r message=FALSE, warning=FALSE}
#load data
df <- read.csv("credit-card-cust-behav-data.csv", stringsAsFactors = FALSE, header = TRUE, encoding  = "UTF-8")
```

```{r message=FALSE, warning=FALSE}
str(df)
```

## Describe Data
## Explore Data
## Verify Data Quality


# DATA PREPARATION

## Select Data
## Clean Data
## Construct Data
## Integrate Data
## Format Data


# MODELING

## Select Modeling Technique
## Generate Test Design
## Build Model Parameter Settings
## Assess Model


# EVALUATION

## Evaluate Results
## Approved Models
## Determine Next Steps


# DEPLOYMENT

## Plan Deployment
## Plan Monitoring and Maintenance
## Produce Final Report -------------------------------  KNIT THIS .Rmd FILE WHEN DONE!!!!
## Review Project